Question
========
```{r, echo = FALSE, include = FALSE}
library(haven)
library(dplyr)
starwars <- read_spss("data/starwars.sav") %>%
  select(-c(homeworld,species))
```

Consider the following tibble, how can we cancel the labels of the variable `hair_color`?

```{r, echo = FALSE}
starwars
```


Answerlist
----------
* `starwars$hair_color <- unlabelled(starwars$hair_color)`
* `unlabelled(starwars$hair_color)`
* `val_label(starwars)$hair_color <- NULL`
* `val_labels(starwars)$hair_color <- NULL`
* `var_labels(starwars)$hair_color <- NULL`
* `val_labels(starwars$hair_color) <- NULL`
* `unlabelled(starwars)$hair_color`
* `starwars$hair_color <- unlabelled(starwars)`
* `starwars <- unlabelled(starwars$hair_color)`
* It's not possible to cancel the values labels of only one variable

Solution
========
The correct way to cancel the labels of the variable `hair_color` is `unlabelled(starwars$hair_color)`, and then we have to assign it to `starwars$hair_color`

Answerlist
----------
* Correct
* Wrong: We are not assigning the unlabeled values to the data frame
* Wrong: `val_label()` is used to get the list of value label for a variable, not to set labels
* Wrong: `val_labels()` is used to get the list of value label, not to set labels
* Wrong: `var_labels()` is used to get the list of variable labels, not to set labels
* Wrong: `val_labels()` is used to get the list of value label, not to set labels
* Wrong:  We are not assigning the unlabeled values to the data frame
* Wrong: We are assigning the whole unlabeled data frame to just one variable
* Wrong: We are assigning just one variable to the whole unlabeled data frame
* Wrong: It is possible

Meta-information
================
exname: Label SPSS
extype: schoice
exsolution: 1000000000
exshuffle: 5
