---
title: "Excel Files"
---

```{r, echo=FALSE}
dataDir <- "data/"
```


## Loading Excel Files

Many datasets, especially those shared outside of academic or statistical contexts, come in Excel format. R provides multiple packages to work with Excel workbooks.

The `readxl` package, part of the tidyverse, is a lightweight and reliable tool for reading .xlsx files without requiring external dependencies. We use `read_excel()` to read a specific worksheet:

```{r}
library(readxl)
titanic2 <- read_excel(paste0(dataDir, "example_data.xlsx"), sheet = "titanic")
```

An alternative is the `openxlsx` package, which offers both reading and writing capabilities. It's useful when you need more control over formatting or workbook structure:

```{r}
library(openxlsx)
titanic3 <- read.xlsx(paste0(dataDir, "example_data.xlsx"), sheet = "titanic")
```

After loading the data, we can inspect the structure with:

```{r}
str(titanic2)
str(titanic3)
```

Each function achieves a similar result—reading data from the "titanic" worksheet—but when we want to be more specific about what we want to load, the two functions use different command. `readxl` is more minimal, while `openxlsx` is more feature-rich.

